### 2022-07-31 analysis.Rmd
update data

---
title: "Illusion Game Validation"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3)

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

## Introduction

Behavioural findings regarding the Illusion Game.

## Methods

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

# This is a local folder containing raw data from unzipped pavlovia
dfraw <- read.csv("data/study1.csv") |> 
  mutate(Screen_Refresh = as.character(Screen_Refresh),
         Illusion_Side = as.factor(Illusion_Side))
```

### Exclusions

For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).

```{r message=FALSE, warning=FALSE}
dfsub <- dfraw |>
  group_by(Participant) |>
  summarize(Error = sum(Error) / n(),
            RT = mean(RT)) |>
  ungroup()

dfraw |>
  group_by(Participant, Illusion_Type, Block) |>
  summarize(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |>
  estimate_density(at = c("Illusion_Type", "Block")) |>
  ggplot(aes(x = x, y = y)) +
  geom_line(aes(color = Illusion_Type, linetype= as.factor(Block)))

df <- dfraw |>
  group_by(Participant, Illusion_Type, Block) |>
  mutate(ErrorRate_per_block = sum(Error) / n()) |> 
  ungroup() |> 
  filter(ErrorRate_per_block < 0.5) |> 
  select(-ErrorRate_per_block)

nrow(dfraw) - nrow(df)
```

### Participants

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |> 
  slice(1) |> 
  ungroup()
```

`r report::report_participants(dfsub, age="Age", sex="Sex")`

```{r }
plot_distribution <- function(dfsub, what="Age", title=what, subtitle="", fill="orange") {
  dfsub |> 
    ggplot(aes_string(x = what)) +
    geom_density(fill = fill) + 
    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype="dashed") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(title, subtitle=subtitle) +
    theme_modern() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5),
          plot.subtitle = element_text(face = "italic", hjust = 0.5),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank())
}

plot_waffle <- function(data, title="Nationality") {
  data |> 
    # mutate(label = emojifont::fontawesome('fa-twitter')) |> 
    ggplot(aes(x, y, fill = group)) + 
    ggwaffle::geom_waffle() +
    # geom_point() +
    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
    coord_equal() + 
    ggtitle(title) +
    labs(fill = "") +
    theme_void() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```

```{r }
p1 <- plot_distribution(dfsub, "Age", fill="#FF9800")
p2 <- plot_distribution(dfsub, "Duration", title="Total Duration", subtitle="in minutes", fill="#F44336")
p3 <- plot_distribution(dfsub, "Break_Duration", title="Break Duration", subtitle="in minutes", fill="#3F51B5")

p4 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Sex)) |> 
  plot_waffle("Sex") +
  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))

p5 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Education)) |> 
  plot_waffle("Education") +
  scale_fill_viridis_d()

p6 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Nationality)) |> 
  plot_waffle("Nationality") +
  scale_fill_viridis_d()

p7 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Ethnicity)) |> 
  plot_waffle("Ethnicity") +
  scale_fill_viridis_d()

p8 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Resolution)) |> 
  plot_waffle("Screen Resolution") +
  scale_fill_viridis_d()

p9 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Device_OS)) |> 
  plot_waffle("Device OS") +
  scale_fill_viridis_d()

# p10 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Refresh)) |> 
#   plot_waffle("Screen Refresh Rate") +
#   scale_fill_viridis_d()

(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9) 
```

## Results

### Manipulation Checks

We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1|Participant), 
                          data = filter(df, Illusion_Strength == 0), 
                          family="binomial")

parameters(model, effects="fixed") |> 
  tail(10)
```

### Full Models

#### Outliers

```{r message=FALSE, warning=FALSE}
p1 <- estimate_density(df, select="RT", at="Participant") |> 
  normalize(select = "y") |> 
  ggplot(aes(x=x, y=y)) +
  geom_area(data=normalize(estimate_density(df, select="RT"), select="y"), alpha=0.2) +
  geom_line(aes(color=Participant, group=Participant)) + 
  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
  scale_color_material_d("rainbow", guide="none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, 5000)) +
  theme_modern() +
  theme(axis.text.y = element_blank())

df$Outlier <- df$RT > 2500

p2 <- df |> 
  group_by(Participant) |> 
  summarize(Outlier = sum(Outlier) / n()) |> 
  mutate(Participant = fct_reorder(Participant, Outlier)) |> 
  ggplot(aes(x = Participant, y = Outlier)) +
  geom_bar(stat="identity", aes(fill=Participant)) +
  scale_fill_material_d(guide="none") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
  theme_modern("rainbow") +
  theme(axis.text.x = element_blank())

p1 | p2

df <- filter(df, Outlier == FALSE)
```

#### Errors

```{r message=FALSE, warning=FALSE}
model_errors <- function(df, illusion = "Delboeuf") {
  data <- filter(df, Illusion_Type == illusion)

  # formula <- brms::bf(
  #   # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
  #   Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
  #   family = "bernoulli"
  # )
  # model <- brms::brm(formula,
  #   data = data,
  #   algorithm = "meanfield",
  #   refresh = 0
  # )
  # model <- glmmTMB::glmmTMB(
  #   Error ~ Illusion_Difference * Illusion_Strength + (1 | Participant),
  #   data = data,
  #   family = "binomial"
  # )
  model <- mgcv::gamm(
    Error ~ s(Illusion_Difference, by = Illusion_Strength, k=4),
    random = list(Participant = ~1),
    data = data,
    family = "binomial"
  )

  n_lines <- 5
  pred <- estimate_relation(model, at = c("Illusion_Strength", "Illusion_Difference"), length = c(30, n_lines)) |>
    mutate(Illusion_Difference = as.factor(Illusion_Difference),
           Illusion_Type = illusion)

  p <- pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted, group = Illusion_Difference)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept=c(0.05, 0.5, 0.95), linetype="dotted", alpha=0.5) +
    geom_line(aes(color = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)) +
    scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)) +
    theme_modern() +
    labs(
      title = paste0(illusion, " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  list(p = p, model = model, pred = pred)
}
```

```{r message=FALSE, warning=FALSE}
rez_ponzo <- model_errors(df, illusion = "Ponzo")
rez_mullerlyer <- model_errors(df, illusion = "Müller-Lyer")
rez_delboeuf <- model_errors(df, illusion = "Delboeuf")
rez_ebbinghaus <- model_errors(df, illusion = "Ebbinghaus")
rez_contrast <- model_errors(df, illusion = "Contrast")
rez_white <- model_errors(df, illusion = "White")
rez_poggendorff <- model_errors(df, illusion = "Poggendorff")
rez_zollner <- model_errors(df, illusion = "Zöllner")
rez_rodframe <- model_errors(df, illusion = "Rod-Frame")
rez_verticalhorizontal <- model_errors(df, illusion = "Vertical-Horizontal")
```

```{r message=FALSE, warning=FALSE}
rez_ponzo$p
rez_mullerlyer$p
rez_delboeuf$p
rez_ebbinghaus$p
rez_verticalhorizontal$p
rez_rodframe$p
rez_zollner$p
rez_poggendorff$p
rez_contrast$p
rez_white$p
```

<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- rbind( -->
<!--   rez_delboeuf$pred, -->
<!--   rez_ebbinghaus$pred, -->
<!--   rez_ponzo$pred, -->
<!--   rez_mullerlyer$pred, -->
<!--   rez_contrast$pred, -->
<!--   rez_white$pred, -->
<!--   rez_poggendorff$pred, -->
<!--   rez_zollner$pred, -->
<!--   rez_rodframe$pred, -->
<!--   rez_verticalhorizontal$pred -->
<!--   ) |>  -->
<!--   ggplot(aes(x = Illusion_Strength, y = Predicted, group = Illusion_Difference)) + -->
<!--   geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) + -->
<!--   # geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_line(aes(color = Illusion_Difference)) + -->
<!--   scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) + -->
<!--   scale_x_continuous(expand = c(0, 0)) + -->
<!--   scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--   scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--   theme_modern() + -->
<!--   labs( -->
<!--     title = paste0(illusion, " Illusion"), -->
<!--     color = "Difficulty", fill = "Difficulty", -->
<!--     y = "Probability of Error", -->
<!--     x = "Illusion Strength" -->
<!--   ) + -->
<!--   theme(plot.title = element_text(face = "bold", hjust = 0.5)) + -->
<!--   facet_wrap(~Illusion_Type, scales="free") -->
<!-- ``` -->

#### RT

<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- run_model_rt <- function(df, illusion = "Delboeuf") { -->
<!--   data <- df |> -->
<!--     filter(Illusion_Type == illusion) |>  -->
<!--     mutate(Error = ifelse(Error == 1, "Error", "Correct")) -->

<!--   model <- lme4::lmer( -->
<!--     RT ~ Error * (Illusion_Difference * Illusion_Strength) + (1 | Participant), -->
<!--     data = data -->
<!--   ) -->

<!--   pred <- estimate_relation(model, at = c("Illusion_Strength", "Illusion_Difference", "Error"), length = c(30, 5)) |> -->
<!--     mutate(Illusion_Difference = as.factor(Illusion_Difference), -->
<!--            Illusion_Type = illusion) -->

<!--   p <- pred |> -->
<!--     mutate(Predicted = ifelse(Error == "Error", -Predicted, Predicted)) |>  -->
<!--     ggplot(aes(x = Illusion_Strength, y = Predicted, group = interaction(Illusion_Difference, Error))) + -->
<!--     # geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) + -->
<!--     geom_line(aes(color = Illusion_Difference)) + -->
<!--     scale_y_continuous(expand = c(0, 0)) + -->
<!--     scale_x_continuous(expand = c(0, 0)) + -->
<!--     scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--     scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--     theme_modern() + -->
<!--     labs( -->
<!--       title = paste0(illusion, " Illusion"), -->
<!--       color = "Difficulty", fill = "Difficulty", -->
<!--       y = "Reaction Time", -->
<!--       x = "Illusion Strength" -->
<!--     ) + -->
<!--     theme(plot.title = element_text(face = "bold", hjust = 0.5)) -->

<!--   list(p = p, model = model, pred = pred) -->

<!-- } -->
<!-- ``` -->
<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- run_model_rt(df, illusion = "Delboeuf")$p -->
<!-- run_model_rt(df, illusion = "Ebbinghaus")$p -->
<!-- run_model_rt(df, illusion = "Ponzo")$p -->
<!-- run_model_rt(df, illusion = "Müller-Lyer")$p -->
<!-- run_model_rt(df, illusion = "Contrast")$p -->
<!-- run_model_rt(df, illusion = "White")$p -->
<!-- run_model_rt(df, illusion = "Poggendorff")$p -->
<!-- run_model_rt(df, illusion = "Zöllner")$p -->
<!-- run_model_rt(df, illusion = "Rod-Frame")$p -->
<!-- run_model_rt(df, illusion = "Vertical-Horizontal")$p -->
<!-- ``` -->

### Inter-Illusion Parameters Standardization

### Individual Score Approximation

correlation between inter-individual indices extracted from full model with that of individual models (with and without priors)

<!-- ### Drift Model -->

## References