### 2022-08-05 study0/analysis.Rmd
update stim

---
title: "Illusion Game Validation (Study 1)"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3,
        mc.cores = 4,
        brms.algorithm = "meanfield",
        brms.backend = "cmdstanr")

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

source("preprocessing.R")

cbrt <- function(x) x^(1/3)

df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
df$Illusion_Difference_cbrt <- cbrt(df$Illusion_Difference)

df <- filter(df, RT < 900)
```

## Results {.tabset}

### Delboeuf

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Delboeuf")
```

```{r message=FALSE, warning=FALSE}
best_models <- function(data) {
  models_err <- list()
  models_rt <- list()
  for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
    for(side in c("", "-side")) {
      name <- paste0("dif", k1, side)
      print(name)
      f <- paste0("Illusion_Difference", 
                  k1,
                  " + (1|Participant)"
                  )

      if(side == "-side") f <- paste0("Illusion_Side * ", f)

      m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)), 
                            data = data, family = "binomial")
      if(performance::check_convergence(m)) {
        models_err[[name]] <- m
      }
      m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)), 
                            data = data)
      if(performance::check_convergence(m)) {
        models_rt[[name]] <- m
      }
    }
  }

  to_keep <- compare_performance(models_err, metrics = c("BIC")) |> 
    arrange(BIC) |> 
    slice(1:5) |> 
    pull(Name)

  test <- test_performance(models_err[to_keep], reference=1)
  perf <- compare_performance(models_err[to_keep], metrics = c("BIC", "R2"))

  rez <- merge(perf, test) |> 
    arrange(BIC) |> 
    select(Name, BIC, R2_marginal, BF) |> 
    mutate(BF = insight::format_bf(BF, name=""),
           Model = "Error")

  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |> 
    arrange(BIC) |> 
    slice(1:5) |> 
    pull(Name)

  test <- test_performance(models_rt[to_keep], reference=1)
  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))

  merge(perf, test) |> 
    arrange(BIC) |> 
    select(Name, BIC, R2_marginal, BF) |> 
    mutate(BF = insight::format_bf(BF, name=""),
           Model = "RT") |> 
    rbind(rez)
}

best_models(data)
```

```{r message=FALSE, warning=FALSE}
visualize <- function(data, model_err, model_rt) {

  at <- list(Illusion_Difference=seq(0, max(data$Illusion_Difference), length.out=30),
             Illusion_Side = levels(unique(data$Illusion_Side)))

  model <- mgcv::gamm(Error ~ s(Illusion_Difference, by = Illusion_Side),
                      random = list(Participant = ~1) ,
                      data = data,
                      family = "binomial")
  model_lin <- glmmTMB::glmmTMB(Error ~ Illusion_Difference + (1|Participant) ,
                      data = data,
                      family = "binomial")

  p1 <- data |> 
    group_by(Illusion_Difference, Illusion_Side) |> 
    summarize(Error = mean(Error)) |> 
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(stat="identity") +
    geom_hline(yintercept=0.5, linetype="dotted") +
    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_err, at=at[1]), aes(y=Predicted), color="red") +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  model <- mgcv::gamm(RT ~ s(Illusion_Difference, by = Illusion_Side),
                     random = list(Participant = ~1) ,
                     data = filter(data, Error == 0, RT < 1000))
    model_lin <- glmmTMB::glmmTMB(RT ~ Illusion_Difference + (1|Participant) ,
                      data = data)

  p2 <- data |> 
    filter(RT < 1000) |> 
    # group_by(Illusion_Difference, Illusion_Side) |> 
    # summarize(RT = mean(RT)) |> 
    ggplot(aes(x = Illusion_Difference, y = RT)) +
    ggdist::stat_pointinterval(aes(group=Illusion_Difference)) +
    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_rt, at=at[1]), aes(y=Predicted), color="red") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "RT",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  p1 / p2
}

model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

```{r message=FALSE, warning=FALSE}
optimal <- function(model_err) {
  pred <- estimate_relation(model_err, length=500)

  bounds <- pred |> 
      slice(c(which.min(abs(Predicted - 0.25)), 
              which.min(abs(Predicted - 0.025))))

  rez <- bounds |> 
    transmute(Illusion_Difference_log = Illusion_Difference_log,
              Illusion_Difference = exp(Illusion_Difference_log) - 1,
              Error = insight::format_value(Predicted, as_percent=TRUE))

  p <- pred |> 
    ggplot(aes(y = Predicted, x = Illusion_Difference_log)) +
    geom_line() +
    geom_hline(yintercept = c(0.25, 0.025), linetype="dotted", color=c("red", "green")) +
    geom_vline(xintercept = bounds$Illusion_Difference_log, linetype="dotted", color=c("red", "green")) +
    theme_modern()
  list(rez=rez, p=p)
}

model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
                              data = data, family = "binomial")

rez <- optimal(model_err)
rez$rez
rez$p
```

### Müller-Lyer

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Müller-Lyer")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
                              data = data, family = "binomial")

rez <- optimal(model_err)
rez$rez
rez$p
```

### Vertical-Horizontal

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Vertical-Horizontal")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ sqrt(Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### Poggendorff

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Poggendorff")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### Zöllner

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Zöllner")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### Rod-Frame

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Rod-Frame")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### Contrast

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Contrast")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### White

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "White")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```