### 2022-08-05 study2.Rmd
update RodFrame stim

---
title: "Illusion Game Validation (Study 2)"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3,
        mc.cores = 4,
        brms.algorithm = "sampling",
        brms.backend = "cmdstanr")

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

# Introduction

The goal of study 2 was to...

# Methods

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# PREPROCESSING ================================================================
source("preprocessing.R")

# This is a local folder containing raw data from unzipped pavlovia
# It has been added to .gitignore to NOT be published on github
# (it contains the subject ID of the participants)
participants <- list.files("study2/data/")

df <- data.frame()
for (ppt in participants) {
  df <- rbind(df, preprocess_raw(file = paste0("study2/data/", ppt)))
}

df$Pyllusion <- "1.2"

# Fix 
df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"])

# Transformation
df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))

# Save anonymized data
write.csv(df, "data/study2.csv", row.names = FALSE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

df <- read.csv("data/study2.csv") |>
  mutate(
    Date = as.Date(Date),
    Participant = fct_reorder(Participant, Date),
    Screen_Refresh = as.character(Screen_Refresh),
    Illusion_Side = as.factor(Illusion_Side),
    Block = as.factor(Block)
    # Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
  )
```

## Exclusions {.tabset}

```{r message=FALSE, warning=FALSE}
# plot(estimate_density(dfraw[dfraw$Participant == "60684f29dbfe1bb2059e5e27_rkqoy", "RT"]))

# Dear participant, thank you for participating in our study. Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable. We understand that you might have been in a hurry or had some other issues; we hope to open-up more slots in the future would you be interested to participate again.

outliers <- c(
  # Error rate of 48.8% Very short RT
  # Prolific Status: NA
  "5e66ed53de7896486f9a71f8_p8ckk"
  )
```

We removed `r length(outliers)` participants upon inspection of the average error rage (when close to 50%, suggesting random answers) and/or when the reaction time distribution was implausibly fast.

For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).

### Error Rate

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  summarize(
    # n = n(),
    Error = sum(Error) / n(),
    RT_Mean = mean(RT),
    RT_SD = sd(RT),
  ) |>
  ungroup() |>
  arrange(desc(Error))

knitr::kable(dfsub) |>
  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background = "#EF9A9A")
```

### Reaction Time Distribution

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
# RT distribution
p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
  group_by(Participant) |>
  normalize(select = "y") |>
  ungroup() |>
  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
  ggplot(aes(x = x, y = y)) +
  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
  geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
  geom_vline(xintercept = 150, linetype = "dashed", color = "red") +
  scale_color_manual(values = c("red" = "red", "blue" = "blue"), guide = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, 3000)) +
  theme_modern() +
  theme(axis.text.y = element_blank()) +
  facet_wrap(~Participant) +
  labs(y = "", x = "Reaction Time (ms)")
# p
# ggsave("figures/outliers.png", p, width=20, height=15)

# Filter out
df <- filter(df, !Participant %in% outliers)
```

### Error Rate per Illusion Block

```{r message=FALSE, warning=FALSE}
temp <- df |>
  group_by(Participant, Illusion_Type, Block) |>
  summarize(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |>
  arrange(desc(ErrorRate_per_block))

temp2 <- temp |>
  filter(ErrorRate_per_block >= 0.5) |>
  group_by(Illusion_Type, Block) |>
  summarize(n = n()) |>
  arrange(desc(n), Illusion_Type) |>
  ungroup() |>
  mutate(
    n_trials = cumsum(n * 56),
    p_trials = n_trials / nrow(df)
  )

# knitr::kable(temp2)

p1 <- temp |>
  estimate_density(at = c("Illusion_Type", "Block")) |>
  ggplot(aes(x = x, y = y)) +
  geom_line(aes(color = Illusion_Type, linetype = Block)) +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(y = "Distribution", x = "Error Rate") +
  theme_modern()

p2 <- temp2 |>
  mutate(Block = fct_rev(Block)) |>
  ggplot(aes(x = Illusion_Type, y = p_trials)) +
  geom_bar(stat = "identity", aes(fill = Block)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  labs(y = "Percentage of Trials Removed", x = "Illusion Type") +
  theme_modern() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p1 | p2

# Drop
df <- df |>
  group_by(Participant, Illusion_Type, Block) |>
  mutate(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |>
  filter(ErrorRate_per_block < 0.5) |>
  select(-ErrorRate_per_block)

rm(temp, temp2)
```

### Reaction Time per Trial

```{r message=FALSE, warning=FALSE}
df <- df |> 
  group_by(Participant, Error) |> 
  mutate(Outlier = ifelse(Error == 0 & (RT < 150 | standardize(RT) > 4), TRUE, FALSE)) |> 
  ungroup()

p1 <- estimate_density(df, select = "RT", at = "Participant") |>
  group_by(Participant) |>
  normalize(select = "y") |>
  ungroup() |>
  merge(df |> 
    filter(Error == 0) |> 
    group_by(Participant) |> 
    summarize(Outlier = mean(RT) + 4 * sd(RT))) |> 
  mutate(Outlier = ifelse(x >= Outlier, TRUE, FALSE)) |> 
  ggplot(aes(x = x, y = y)) +
  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
  geom_line(aes(color = Participant, linetype=Outlier)) +
  geom_vline(xintercept = c(150), linetype = "dashed", color = "red") +
  scale_color_material_d("rainbow", guide = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  guides(linetype = "none") +
  coord_cartesian(xlim = c(0, 4000)) +
  theme_modern() +
  theme(axis.text.y = element_blank()) +
  labs(y = "", x = "Reaction Time (ms)")

p2 <- df |>
  group_by(Participant) |>
  summarize(Outlier = sum(Outlier) / n()) |>
  mutate(Participant = fct_reorder(Participant, Outlier)) |>
  ggplot(aes(x = Participant, y = Outlier)) +
  geom_bar(stat = "identity", aes(fill = Participant)) +
  scale_fill_material_d("rainbow", guide = "none") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
  see::theme_modern() +
  theme(axis.text.x = element_blank())

p1 | p2
```

We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (150 ms < RT < 4 SD above mean).

```{r message=FALSE, warning=FALSE}
df <- filter(df, Outlier == FALSE)
```

## Participants

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |>
  slice(1) |>
  ungroup()
```

The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.

```{r }
plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
  dfsub |>
    ggplot(aes_string(x = what)) +
    geom_density(fill = fill) +
    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(title, subtitle = subtitle) +
    theme_modern() +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      plot.subtitle = element_text(face = "italic", hjust = 0.5),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    )
}

plot_waffle <- function(dfsub, what = "Nationality", title = what) {
  ggwaffle::waffle_iron(dfsub, what) |>
    # mutate(label = emojifont::fontawesome('fa-twitter')) |>
    ggplot(aes(x, y, fill = group)) +
    ggwaffle::geom_waffle() +
    # geom_point() +
    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
    coord_equal() +
    ggtitle(what) +
    labs(fill = title) +
    theme_void() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```

```{r fig.width=20, fig.height=15}
p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")

p4 <- plot_waffle(dfsub, "Sex") +
  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))

p5 <- plot_waffle(dfsub, "Education") +
  scale_fill_viridis_d()

p6 <- plot_waffle(dfsub, "Nationality") +
  scale_fill_metro_d()

p7 <- plot_waffle(dfsub, "Ethnicity") +
  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))

p8 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
  scale_fill_pizza_d()

p9 <- plot_waffle(dfsub, "Device_OS", title = "Device OS") +
  scale_fill_bluebrown_d()

# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
#   scale_fill_viridis_d()

(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9)
```

## Models {.tabset}

### Delboeuf

#### Model Selection

```{r message=FALSE, warning=FALSE, cache=cache}
best_models <- function(data) {
  models_err <- list()
  models_rt <- list()
  for (k1 in c("", "_log", "_sqrt", "_cbrt")) {
    for (k2 in c("", "_log", "_sqrt", "_cbrt")) {
      for (side in c("", "--SIDE")) {
        name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side)
        message(name)
        f <- paste0(
          "Illusion_Difference",
          k1,
          " * Illusion_Strength",
          k2,
          " + (1|Participant)"
        )

        if (side == "--SIDE") f <- paste0("Illusion_Side * ", f)

        m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
          data = data, family = "binomial"
        )
        if (performance::check_convergence(m)) {
          models_err[[name]] <- m
        }

        m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)),
          data = data
        )
        if (performance::check_convergence(m)) {
          models_rt[[name]] <- m
        }
      }
    }
  }

  # RT
  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |>
    arrange(BIC) |>
    slice(1:5) |>
    pull(Name)

  test <- test_performance(models_rt[to_keep], reference = 1)
  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))
  rt <- merge(perf, test) |>
    arrange(BIC) |>
    select(Name, BIC, R2_marginal, BF) |>
    mutate(
      BF = insight::format_bf(BF, name = ""),
      Model = "RT"
    )

  # Errors
  to_keep <- compare_performance(models_err, metrics = c("BIC")) |>
    arrange(BIC) |>
    slice(1:5) |>
    pull(Name)

  test <- test_performance(models_err[to_keep], reference = 1)
  perf <- compare_performance(models_err[to_keep], metrics = c("BIC", "R2"))

  merge(perf, test) |>
    arrange(BIC) |>
    select(Name, BIC, R2_marginal, BF) |>
    mutate(
      BF = insight::format_bf(BF, name = ""),
      Model = "Errors"
    ) |>
    rbind(rt)
}

data <- filter(df, Illusion_Type == "Delboeuf")

best_models(data)
```

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
plot_descriptive_err <- function(data, side = "leftright") {
  # Sanity checks
  if (length(unique(data$Illusion_Difference)) != 8) stop("Illusion_Difference values != 8")
  if (length(unique(data$Illusion_Strength)) != 15) stop("Illusion_Strength values != 15")

  if (side == "leftright") {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if (x == "Left") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
    } else if (x == "Right") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
    }
  } else {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if (x == "Up") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
    } else if (x == "Down") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
    }
    data$Answer <- fct_rev(data$Answer)
  }

  dodge1 <- 0.1 * diff(range(data$Illusion_Difference))
  dodge2 <- -0.1 * diff(range(data$Illusion_Strength))

  p1 <- data |>
    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
    summarize(Error = mean(Error), .groups = "drop") |>
    ungroup() |>
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width=0.03) +
    # geom_line(aes(color = Illusion_Strength, group=Illusion_Strength), position = position_dodge(width=dodge1)) +
    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
    theme_modern() +
    labs(
      color = "Illusion Strength",
      fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  p2 <- data |>
    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
    summarize(Error = mean(Error), .groups = "drop") |>
    ungroup() |>
    # mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 4))) |>
    ggplot(aes(x = Illusion_Strength, y = Error)) +
    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width=0.1) +
    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    theme_modern() +
    labs(
      color = "Task Difficulty",
      fill = "Task Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  if (side == "leftright") {
    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
      plot_annotation(
        title = paste(data$Illusion_Type[1], "Illusion"),
        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
      )
  } else {
    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
      plot_annotation(
        title = paste(data$Illusion_Type[1], "Illusion"),
        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
      )
  }
  p
}

plot_descriptive_err(data)
```

##### Model Specification

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Difference * Illusion_Strength + 
    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
  family = "bernoulli"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

model_delboeuf_err <- brms::brm(formula,
  data = data,
  refresh = 0
)

# parameters::parameters(model)
```

### Model Visualization

```{r message=FALSE, warning=FALSE, cache=cache}
plot_model_err <- function(data, model) {
  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))

  # Get variables
  vars <- insight::find_predictors(model)$conditional
  vardiff <- vars[1]
  varstrength <- vars[2]

  # Get predicted
  pred <- estimate_relation(model,
                            at = vars,
                            length = c(NA, 25))

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
  names(colors) <- diffvals

  # Assign color from the same palette to every observation of data (for geom_dots)
  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
  data$color <- colors[as.character(closest)]
  data$color <- fct_reorder(data$color, closest)

  # Manual jittering
  xrange <- 0.05*diff(range(data[[varstrength]]))
  data$x <- data[[varstrength]]
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  pred |>
    ggplot(aes_string(x = varstrength, y = "Predicted")) +
    geom_dots(
      data = data,
      aes(x=x, 
          y = Error, 
          group = Error,
          side = .dots_side, 
          # order=as.numeric(color) * ifelse(Error == 1, 1, -1)), 
          order=color), 
      fill = data$color,
      color = NA, 
      alpha=0.5) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes_string(color = vardiff, group = vardiff)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
    theme_modern() +
    labs(
      title = paste0(data$Illusion_Type[1], " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

plot_model_err(data, model_delboeuf_err)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
plot_descriptive_rt <- function(data, side = "leftright") {
  if (side == "leftright") {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if (x == "Left") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
    } else if (x == "Right") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
    }
  } else {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if (x == "Up") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
    } else if (x == "Down") {
      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
    }
    data$Answer <- fct_rev(data$Answer)
  }

  p1 <- data |>
    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
    summarize(RT = mean(RT), .groups = "drop") |>
    ggplot(aes(x = Illusion_Difference, y = RT)) +
    ggdist::stat_slabinterval(aes(fill = Illusion_Strength, group = Illusion_Strength, color = Illusion_Strength), alpha = 0.5, normalize = "groups") +
    # geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
    theme_modern() +
    labs(
      color = "Illusion Strength",
      fill = "Illusion Strength",
      y = "Reaction Time (ms)",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  p2 <- data |>
    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
    summarize(RT = mean(RT), .groups = "drop") |>
    ggplot(aes(x = Illusion_Strength, y = RT)) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
    ggdist::stat_slabinterval(aes(fill = Illusion_Difference, group = Illusion_Difference, color = Illusion_Difference), alpha = 0.5, normalize = "groups") +
    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    theme_modern() +
    labs(
      color = "Task Difficulty",
      fill = "Task Difficulty",
      y = "Reaction Time (ms)",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  if (side == "leftright") {
    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
      plot_annotation(
        title = paste(data$Illusion_Type[1], "Illusion"),
        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
      )
  } else {
    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
      plot_annotation(
        title = paste(data$Illusion_Type[1], "Illusion"),
        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
      )
  }
  p
}

data <- filter(data, Error == 0)

plot_descriptive_rt(data)
```

##### Model Specification

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Difference * Illusion_Strength + 
    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
  family = "bernoulli"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

model_delboeuf_err <- brms::brm(formula,
  data = data,
  refresh = 0
)

# parameters::parameters(model)
```

### Model Visualization

```{r message=FALSE, warning=FALSE, cache=cache}
plot_model_err <- function(data, model) {
  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))

  # Get variables
  vars <- insight::find_predictors(model)$conditional
  vardiff <- vars[1]
  varstrength <- vars[2]

  # Get predicted
  pred <- estimate_relation(model,
                            at = vars,
                            length = c(NA, 25))

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
  names(colors) <- diffvals

  # Assign color from the same palette to every observation of data (for geom_dots)
  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
  data$color <- colors[as.character(closest)]
  data$color <- fct_reorder(data$color, closest)

  # Manual jittering
  xrange <- 0.05*diff(range(data[[varstrength]]))
  data$x <- data[[varstrength]]
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  pred |>
    ggplot(aes_string(x = varstrength, y = "Predicted")) +
    geom_dots(
      data = data,
      aes(x=x, 
          y = Error, 
          group = Error,
          side = .dots_side, 
          # order=as.numeric(color) * ifelse(Error == 1, 1, -1)), 
          order=color), 
      fill = data$color,
      color = NA, 
      alpha=0.5) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes_string(color = vardiff, group = vardiff)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
    theme_modern() +
    labs(
      title = paste0(data$Illusion_Type[1], " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

plot_model_err(data, model_delboeuf_err)
```

### Ebbinghaus

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Ebbinghaus")

plot_descriptive_err(data)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Ebbinghaus", Error == 0)

plot_descriptive_rt(data)
```

### Rod and Frame

#### Model Selection

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Rod-Frame")

best_models(data)
```

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
plot_descriptive_err(data)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(data, Error == 0)

plot_descriptive_rt(data)
```

### Vertical-Horizontal

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Vertical-Horizontal")

plot_descriptive_err(data)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0)

plot_descriptive_rt(data)
```

### Zöllner

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Zöllner")

plot_descriptive_err(data)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Zöllner", Error == 0)

plot_descriptive_rt(data)
```

### White

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "White")

plot_descriptive_err(data)
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "White", Error == 0)

plot_descriptive_rt(data)
```

### Müller-Lyer

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Müller-Lyer")

plot_descriptive_err(data, side = "updown")
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Müller-Lyer", Error == 0)

plot_descriptive_rt(data, side = "updown")
```

### Ponzo

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Ponzo")

plot_descriptive_err(data, side = "updown")
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Ponzo", Error == 0)

plot_descriptive_rt(data, side = "updown")
```

### Poggendorff

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Poggendorff")

plot_descriptive_err(data, side = "updown")
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Poggendorff", Error == 0)

plot_descriptive_rt(data, side = "updown")
```

### Contrast

#### Error Rate

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Contrast")

plot_descriptive_err(data, side = "updown")
```

#### Reaction Time

##### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Contrast", Error == 0)

plot_descriptive_rt(data, side = "updown")
```

## Individual Scores

### Delboeuf

#### Reliability

```{r message=FALSE, warning=FALSE, cache=cache}
random <- model_delboeuf_err 
random <- as.data.frame(model_delboeuf_err)
random <- random[str_detect(names(random), "Participant\\[")]
random <- random |> 
  mutate(Draw = 1:nrow(random)) |>
  pivot_longer(-Draw, names_to="Parameter", values_to = "Value") |> 
  separate(Parameter, into = c("Participant", "Parameter"), sep = ",") |> 
  mutate(Participant = str_remove(Participant, "r_Participant\\["),
         Parameter = str_remove(Parameter, "]"))

random |> 
  ggplot(aes(y= Participant, x = Value)) +
  ggdist::stat_slabinterval() +
  facet_wrap(~Parameter)

data <- random |> 
  mutate(Parameter = paste0("Delboeuf_", Parameter)) |> 
  group_by(Participant, Parameter) |> 
  summarize(Score = as.numeric(bayestestR::map_estimate(Value)), .groups="drop") |>
  ungroup() |> 
  pivot_wider(names_from = Parameter, values_from = Score)

```

#### Approximation

```{r message=FALSE, warning=FALSE, cache=cache}
dat <- data.frame()
for(participant in unique(data$Participant)) {
  freq <- glm(Error ~ Illusion_Difference * Illusion_Strength, 
              family = "binomial",
              data=filter(df, Participant == participant, 
                          Illusion_Type == "Delboeuf"))

  parameters::parameters(freq)[1:2] |> 
    as.data.frame() |> 
    mutate(Participant = participant, 
           Parameter = paste0("freq_Delboeuf_", Parameter))
}
```

## Structure

## Variability